### 2018-01-11 paper/paper.tex
eric comments incorporated

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{listings}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\makeatletter
\newcommand\fs@norules{\def\@fs@cfont{\bfseries}\let\@fs@capt\floatc@ruled
  \def\@fs@pre{}%
  \def\@fs@post{}%
  \def\@fs@mid{\kern3pt}%
  \let\@fs@iftopcapt\iftrue}
\makeatother
\floatstyle{norules}
\restylefloat{algorithm}

\let\tinymatrix\smallmatrix
\let\endtinymatrix\endsmallmatrix
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\scriptstyle}{\scriptscriptstyle}{}{}
\patchcmd{\tinymatrix}{\vcenter}{\vtop}{}{}
\patchcmd{\tinymatrix}{\bgroup}{\bgroup\scriptsize}{}{}

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\usepackage[draft,index]{fixme}
\fxsetup{theme=color,mode=multiuser,layout=inline,draft}

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10010940.10010971.10010980.10010986</concept_id>
%<concept_desc>Software and its engineering~Massively parallel systems</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10010147.10010169</concept_id>
%<concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%<concept>
%<concept_id>10010147.10010257.10010282.10010284</concept_id>
%<concept_desc>Computing methodologies~Online learning settings</concept_desc>
%<concept_significance>300</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~Massively parallel systems}
%\ccsdesc[300]{Computing methodologies~Parallel computing methodologies}
%\ccsdesc[300]{Computing methodologies~Online learning settings}

%\keywords{Multi-Armed Bandits, EASY-Backfilling, Parallel Job Scheduling, High
%Performance Computing, Simulation.}

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Online Tuning of EASY-Backfilling using Queue Reordering Policies.}

\author{
\IEEEauthorblockN{Eric Gaussier}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, CNRS, Grenoble INP, LIG\\
  eric.gaussier@imag.fr } \\ 
\IEEEauthorblockN{J\'er\^ome Lelong}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LJK\\
jerome.lelong@imag.fr}  \\ 
\IEEEauthorblockN{Valentin Reis, Denis Trystram}\\
\IEEEauthorblockA{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LIG\\
denis.trystram@imag.fr}   
}
\begin{document}

\begin{abstract}

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing platforms. Despite its
  simplicity, and the guarantee of no job starvation, it could still be
  improved on a per-system basis. Such tuning is difficult because of
  non-linearities in the scheduling process. The study conducted in this paper
  considers an online approach to the automatic tuning of the EASY heuristic
  for HPC platforms. More precisely, we consider the problem of selecting a
  reordering policy for the job queue under several feedback modes. We show via
  a comprehensive experimental validation on actual logs that periodic
  simulation of historical data can be used to recover existing
  \textit{in-hindsight} results that allow to divide the average waiting time
  by almost 2. This results holds even when the simulator results are noisy.
  Moreover, we show that good performances can still be obtained without a
  simulator, under what is called bandit feedback - when we can only observe
  the performance of the algorithm that was picked on the live system.  Indeed,
  a simple multi-armed bandit algorithm can reduce the average waiting
  time by 40\%.

\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

\subsection{Context}

Providing the computational infrastuctures needed to solve actual complex
problems arising in the various fields of modern society (including climate
change, health, green energy or security) is a strategic challenge. The main
pillar to adress this problem is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity
leads to sophisticated specific Exascale platforms, and the objective of the
community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly,
system management software still has a long way to go. The existing job and
resource management software allow to run tens of thousands of jobs on hundreds
of thousands of cores. Such software are based on robust policies with
positive properties such as relative simplicity and prevention of job
starvation. However, there is still room for improvement. Here, we study how to
tune the ordering of the submission queues in the frame of the classical
so-called EASY\footnote{Extensible Argonne Scheduling sYstem
(EASY)}-Backfilling family of heuristics using two new online approaches. One
approach is based on a simulator and the second uses a multi-armed bandit
algorithm.

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set. The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data. During the
last few years, there was an explosion in the number of works at the interface
of the HPC and Big Data fields, dealing with learning algorithms. Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g.,\ job execution
times ~\cite{learningruntimes}, memory requirements, etc.). However, only few studies
evaluate the impact of such techniques on the performances of the resource
manager.

Unlike most existing studies, which consider the learning of specific
parameters of a trace, we provide algorithms that use feedback from the system
in order to alter the scheduling behavior online. Our focus is to learn how the
platform behaves within the usual framework of EASY-Backfilling described in
Section~\ref{sec:rw} . More precisely, we provide strategies for choosing the
most appropriate queue reordering policy to be used for job selection. This
approach is assessed by an extensive resampling-based experimental validation
that uses 7 actual traces.

\subsection{Contributions}

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. In this work, we focus on the
average waiting time of the jobs. Each method aims at choosing the best queue
reordering policy in a fixed search set of semantically diverse options.

The first method, \textbf{full feedback}, continuously uses simulation on data gathered
from the past and current system workload. In order to reflect the
uncertainties in the data, we also evaluate a more realistic ``noisy'' variant in
which simulations are imprecise.

The second method, \textbf{bandit feedback}, is a simulator-less approach based on a
multi-armed bandit algorithm, which derives feedback by measuring system
performance.

Both of these methods are evaluated with a specifically developed simulator,
which is open-sourced ~\cite{ocst}.

Our main results are as follows: The approach based on the simulator provides
very good results, close to the best policy in the search set. In essence, we
are able to recover in an online fashion existing results that reduce average
waiting times by 11\% to 60\% compared to EASY-FCFS. Indeed, we show
that the required sample sizes for policy selection in batch schedulers are
small enough to avoid workload resampling in practice. The variant
based on noisy data, which is closer to the actual conditions of uncertainties on
both jobs and platform, behaves similarly. The alternate approach based on
a multi-armed bandit also provides reasonable results at a much lower cost,
reducing the average waiting times by 8\% to 46\% compared to
EASY-FCFS.

Let us emphasize the main results of this study: While the First-Come,
First-Serve approach to minimizing the waiting times is a common default
strategy, it can definitively be outperformed. We show how to automate selection
of a good alternative policy in two cases (depending on whether a simulator is
available or not), and provide the difference in performance to be expected
between both situations.

\section{Related Work}
\label{sec:rw}

This section presents and discusses the most significant results related to job
scheduling and learning algorithms in this context. Let us start by recalling
some basics results on heuristics in HPC platforms:

Parallel job scheduling is an old studied theoretical
problem (~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865})
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied to the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY ~\cite{easy} and
Conservative ~\cite{Mu'alem:2001:UPW:380314.380315} backfilling.

While Conservative Backfilling offers many advantages ~\cite{bfchar}, it has a
significant computational overhead and is more complex, which mainly explains
why many of the machines of the top500 ranking ~\cite{top500} rather use a
variant of EASY-Backfilling instead. There is a large body of work seeking to
improve/tune EASY. Indeed, while the basic mechanism is used by some actual
resource and job management software (most notably SLURM ~\cite{SLURMdocSCHED}),
this is seldom done without fine tuning by system administrators.

The original EASY mechanism refers to a First-Come-First-Serve basis.  Several
works explore how to tune EASY by reordering waiting and/or backfilling queues
~\cite{Tsafrir_easypp_2005}, sometimes even in a randomized manner
~\cite{1592720}; some implementations use this method as well
~\cite{Jackson2001}. However, as successful as they may be, these works do not
address the dependency of scheduling metrics on the workload
~\cite{variability}. These studies most often report \textit{post-hoc}
performance since they compare algorithms after the workload is known.

The dynP scheduler ~\cite{streit_selftuning_2002} suggests a systematic method
to tuning these queues. The algorithm requires simulated scheduling runs at
decision time, therefore the time-to-decision of the scheduling is higher than
that of EASY. The adaptation of EASY was studied under a different framework
than ours in ~\cite{talby2005improving}. Other works study the reservation
mechanism ~\cite{srinivasan2002selective}. Online portfolio approaches are also
studied for the cloud setting ~\cite{deng2013exploring}.

\subsection{Data-aware Resource Management.}

There was a recent focus on leveraging the huge amount of data available in
large scale computing platforms to improve system performance. Some
works use collaborative filtering to co-locate tasks in clouds by estimating
application interference ~\cite{7516031}. Some others are closer to the
application level. For instance,  ~\cite{fmodeling} uses binary classification
to distinguish benign memory faults from application errors so as to execute
recovery algorithms.

Several works use similar techniques in the context of HPC, in
particular ~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling ~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures ~\cite{Nissimov2008}. However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in ~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
that directly learns a good scheduling policy from a given policy space.

Existing work ~\cite{jsspp17} takes this approach offline, by
splitting workload data in a training/testing fashion. This work studies
EASY-Backfilling and shows that while the relative performance of some well-known
priority orders for starting jobs differs between workloads, it is relatively
stable over time for a given workload.

\subsection{Multi-Armed Bandits.}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards ~\cite{bubnow}. At every round an action, called an
"arm" in the literature, must be chosen from a fixed set and the corresponding
reward is observed. The goal of a MAB algorithm is to maximize the total reward
obtained in a successive number of rounds. This is essentially achieved using a
combination of explorative actions, which help estimate the quality of each
arm, and of exploitative actions, which leverage the current estimates by using
a good arm. This problem can be addressed using various hypotheses. Most common
are the (original) stochastic setting and the adversarial setting of regret
minimization. See ~\cite{thompson} for the original work on the stochastic
setting and ~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
algorithms. To the best of our knowledge, the work of ~\cite{Banos} is the
earliest for the adversarial setting; see ~\cite{nonstoch} for the
``Exponential-weight'' family of algorithms. We refer to ~\cite{bubnow} for a
comprehensive review of the field. While these algorithms bound the cumulative
difference in loss to the best arm (the regret), they have functional
constraints such as the fact that the rewards should be contained in a range.
The simple heuristic called \textit{Epsilon-Greedy} introduced
in ~\cite{Auer2002} does not have this requirement. For this reason, this
heuristic will be used in this paper and is described in
Subsection~\ref{sub:bandit}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem Setting}
\label{sec:problem_setting}

%This section presents the systems under study and the scheduling problem.  It
%first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
%problem statement.

\subsection{System Description}
\label{sub:sysdesc}

The crucial part of batch scheduling software is the scheduling algorithm that
determines where and when the submitted jobs are executed. The process is as
follows: jobs are submitted by end-users and queued until the scheduler selects
one of them for running. Each job has a provided bound on the execution time
and some resource requirements (number and type of processing units). Then, the
Resource and Job Management System (RJMS) drives the search for the resources required to execute this job. Finally,
the tasks of the job are assigned to the chosen nodes.

In the classical case, the management software needs to execute a set of
concurrent parallel jobs with rigid (known and fixed) resource requirements on
an HPC platform represented by a pool of $m$ identical resources. This is an
on-line problem since the jobs are submitted over time and their
characteristics are only known when they are released.  Below is the brief
description of the characteristics of job
$j$:

\begin{itemize}

   \item Submission date $r_j$ (also called \textit{release date})

   \item Resource requirement $q_j$ (number of processors)

   \item Actual running time $p_j$ (sometimes called \textit{processing time})

   \item Requested running time $\widetilde{p_j}$ (sometimes called
     \textit{walltime}), which is in general an upper bound of $p_j$.

\end{itemize}

The resource requirement $q_j$ of job $j$ is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the users have incentive to
over-estimate the actual values, since jobs may be ``killed'' if they surpass
the provided value.

\subsection{Brief Description of EASY Backfilling}
\label{sub:easy}

The selection of the job to run is performed according to a scheduling policy,
which determines the order the jobs are executed. EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization ~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e.\ every job will
be scheduled at some moment).

EASY-FCFS uses a job queue to select and backfill jobs.  At any
time a scheduling decision is required (i.e.\ job submission or termination),
the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
and starts jobs until it finds a job that can not be started right away. Then,
it makes a reservation for this job at the earliest predictable time and
starts \textit{backfilling} the job queue in FCFS order using any job that
does not delay the unique reservation.

\subsection{Scheduling Objective}
\label{sub:scheduling_objectives}

A system administrator may use one or multiple cost metric(s). Our study of
scheduling performance relies on the waiting times of the jobs, which is one of
the more commonly used objectives. The waiting time of a job is 
 \begin{equation}
   \textbf{Wait}_j =  start_j-r_j
 \end{equation}
where $start_j$ is the starting time of a job.
Like other cost metrics, the waiting time is usually considered in its
\textit{cumulative} version, which means that one seeks to minimize the average
waiting time. It is worth noting that other metrics such as the maximum waiting
time of all the jobs are also worthy of interest. Unfortunately, this criterion
is antagonistic in nature with the average waiting time. Section~\ref{sub:th}
will outline our approach to address the bi-objective aspect of this problem.

\section{Tuning EASY by Reordering and Thresholding the Job Queue}
\label{sec:framework}

This section presents two mechanisms for safely tuning the EASY-Backfilling:
job queue reordering and job thresholding. Together, these two building blocks
make a robust framework for tuning EASY.

\subsection{Reordering}
\label{subsec:policies}

The EASY heuristic uses a job queue to select and backfill jobs. While this job
queue is ordered in FCFS order in the original heuristic, it is possible to
reorder it at will. We focus on a search space of 12 reordering policies.

\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the most commonly used policy ~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First ~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{wait_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
    where $wait_j$ is the time job $j$ has been waiting until the time at which the decision is taken.
    This corresponds the the value of the stretch(or slowdown) where the user runtime estimate
    is used.
  \item SEXP: Smallest Expansion Factor First 
  \item LRF: Largest Ratio $\frac{\widetilde{p_j}}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ \widetilde{p_j} \times q_j$ First (often called "Total Resources")
  \item SAF: Smallest Area First
\end{enumerate}

This search space is designed with the goal of being as semantically diverse as
possible without making any judgement on which policy should perform well in
practice. It does include most policies from related works that we
are aware of. In the following, we denote these policies by $P_i$ with $i = 1
\ldots 12$.

\subsection{Thresholding}
\label{sub:th}

Reordering the job queue means losing the no-starvation guarantee; some
individual jobs therefore can wait an undue amount of time. It is possible to
introduce a thresholding mechanism to prevent this behavior: When a
queued job's waiting time exceeds a fixed threshold $\Theta$, it is
at the head of the queue. We denote by EASY($P,\Theta$) the scheduling
policy that starts and backfill jobs according to the (thresholded) reordering
policy $P$. For the sake of completeness, Algorithm~\ref{alg:EASY} describes
the EASY($P,\Theta$) heuristic.

\begin{algorithm}[]
  \caption{EASY($P,\Theta$) policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \STATE Sort $Q$ according to $P$
    \STATE Move all jobs of $Q$ for which $wait_j > \Theta$ ahead
    of the queue (breaking ties in FCFS order).
    \\ \textit{//Starting jobs until the machine is full}
    \FOR {job $j$ in Q}
    \IF {$j$ can be started given the current system use}
    \STATE Pop $j$ from Q
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{//Backfilling jobs}
    \FOR {job $j'$ in $Q\setminus\{j\}$}
    \IF {$j'$ can be started without delaying the reservation on $j$.}
    \STATE Pop $j'$ from $Q$
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

\section{Online Tuning}
\label{sec:online}

\begin{figure*}[]
  \centering
  \includegraphics[scale=0.6]{figures/variability.pdf}

  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace
  (see Subsection~\ref{sub:traces}) for 7 different policies with $\Theta = 0$. The policy set is
  reduced as not to obstruct the figure with too many colors.}

  \label{fig:variability}
\end{figure*}

Here, we present our policy selection strategies. We will
refer to the period during which a selected policy is applied as the
\textit{policy period} and will denote the length of this period as $\Delta$.
The time interval is thus divided into periods of equal lengths ($\Delta$):
$\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy
period. A new policy is selected at the beginning of each period and applied
during the whole period.

We further assume that there is a certain regularity among periods, i.e.\ that
the distribution of the job submission process does not radically differ
between consecutive periods. This assumption is validated in the study
presented in ~\cite{jsspp17}. It further entails that the behavior of a policy
during the previous periods reflects its behavior on the current one, so that
the selection of a policy can be based on its past behavior. However, there may
be some variability between different periods for certain cost
metrics ~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:variability} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all 7
policies considered. This indicates that when the cost metric is averaged
over different periods, there is a trade-off to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

The selection of a policy is reminiscent of reinforcement learning, where
solving a search and inference problem in a (perhaps restrictive) policy space
is easier than in the space of the original problem. It is important to note,
however, that a pure reinforcement learning approach is difficult to develop in
our context. Indeed, while we have outlined a reduced action space, the state
space to consider is infinite. This is not prohibitive \textit{per se}, as
modern methods ~\cite{aprl} can bypass dimensionality issues via function
approximation. However, these methods rely on large amounts of batch data, and
we are facing an online problem with no pre-existing data\footnote{The
reinforcement learning literature uses the term ``\textit{on-policy}'' for this
setting.}. We rely here on simpler, yet we believe more effective, strategies
to solve this problem.  These strategies are applied online and rely on exact
simulation, noisy simulation and $\epsilon$-greedy bandit exploration.

\subsection{Policy Selection with Exact Simulation}
\label{sub:feedback}

Several simulators have been developed for ``playing'' reordering policies on a
given set of jobs. We rely in this study on a lightweight simulator (see
Section~\ref{sec:experiments}) that can efficiently simulate different
policies. Such simulators are interesting inasmuch as they provide an estimate
of the cost of a given policy on a set of jobs, as described below.

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the
period $\Delta_t$ ($0 \le t $), and $P_i$ ($1 \le i \le 12$) one reordering
policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$
during period $\Delta_t$ is defined as:
%
\begin{equation} \label{eq:cost-exact} w(t;P_i) = \sum_{j=1}^{l(\Delta_t)}
\mbox{Wait}_j^{P_i} \end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of
the waiting time for job $j$ according to policy $P_i$. The estimation of the
cost of a policy over all the periods preceding the current period can then be
defined as:
%
\begin{equation} \label{eq:tot-cost-exact} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w(t;P_i) \end{equation}
%
where $\lambda \in [0,1]$ is a decaying discount factor that can be used to
privilege recent history (\textit{i.e.} recent periods). The above formulation
is general and covers several possible situations : From stationary job
distributions where $\lambda$ can be set to 1 to non-stationary distributions
where the more recent history has to be privileged. As explained in Section
6.1, in our experiments $\lambda$ is set to 1 as the distributions considered
are stationary but the reader has to bear in mind that our framework can be use
on different distributions by resorting to different values of lambda that can
be set by cross-validation.

One then selects the
policy $P$ that minimizes the above cost for the period $\Delta_T$:
%
\begin{equation} \label{eq:select-exact} P = \mbox{argmin}_{P_i, 1 \le i \le
12} w(\rightarrow T;P_i) \end{equation}

The above cost directly corresponds to the cumulative waiting time of the
policy over the preceding periods (more precisely to the cumulative simulated
estimate of the waiting time of the policy over the preceding periods) so that
the length of the period has little impact here. The only bias comes from the
boundary states of the simulation. Indeed, we run simulations from an empty
system and wait for the system to be empty when job submissions cease at the
end of the period (see Figure~\ref{fig:resimu}).

\begin{figure*}[]
  \centering
  \input{dia/_build/algo.tex}

  \caption{Illustration of the online simulation-based strategy. Every week,
  the wokload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-exact} (or~\eqref{eq:tot-cost-noisy} in the Noisy case
  from Section~\ref{sub:noisy}) and the policy for the next week is chosen
  according to~\eqref{eq:select-exact}}.

  \label{fig:resimu}
\end{figure*}

Furthermore, in the context of this study and as detailed in
Section~\ref{sec:experiments}, we rely on averages over several execution
traces in order to obtain reliable estimate of the behavior of different
policies and policy selection strategies. Such traces are typically obtained by
simulation. Using the same simulator for generating the traces and estimating
the cost as defined in Eqs \eqref{eq:cost-exact} and \eqref{eq:tot-cost-exact}
would however be too optimistic and would represent an upper bound on what can
be achieved by a selection strategy based on simulation.

In order to have a more realistic estimate of the behavior of a selection
strategy based on simulation, we introduce noise in the simulator, as described
below.

\subsection{Policy Selection with Noisy Simulation}
\label{sub:noisy}

We take another step to evaluate how the simulation strategy for selecting
policies would work in a batch scheduler. Simulation of real systems can be
imprecise. While existing work reports values as low as 2\% imprecision we take
a conservative approach and consider imprecision up to of 20\%. We randomly
introduce multiplicative noise in the estimate of the policy cost defined by
Eq.~\eqref{eq:cost-exact} by rescaling it, either down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) = \rho_t^i
\sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus
adding a +/- 20\% multiplicative noise on the waiting time estimated by the
simulator.  This value is taken as to be sufficiently larger than values
previously reported, see e.g. ~\cite{dutot:hal-01333471} which reports 2
percent variations of the simulated metric compared to the metric on the real
systems.  The overall cost is then defined in the same way as before, leading
to:
%
\begin{equation} \label{eq:tot-cost-noisy} w_{\mbox{noisy}}(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before the policy that minimizes the above
cost is selected for the period $\Delta_T$.

\begin{equation} \label{eq:select-noisy} P = \mbox{argmin}_{P_i, 1 \le i \le
12} w_{\mbox{noisy}}(\rightarrow T;P_i) \end{equation}

Instead of introducing noise in the output of the simulator when selecting the
policy, one could have introduced it when collecting the execution traces. Our choice
to add the noise after the simulation step is motivated by simplicity. As we
will see in Section~\ref{sec:experiments}, there is little difference between
the two selection methods, which is an important result of our study.

Lastly, as before, the length of the period has little impact here as the cost
metric corresponds to the cumulative simulated waiting time for each policy.

\subsection{Policy Selection with $\epsilon$-greedy Exploration}
\label{sub:bandit}

The previous strategies estimate the cost of all policies over all previous
periods; the best policy according to this cost is then selected for the
current period, the costs of all policies being updated for the next period.
This is interesting as one maintains a complete knowledge of all policies over
time. However, this requires computing many estimates at each period (as many
as there are policies), which can be time consuming or cumbersome even with
lightweight simulators.

Thus, we explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
~\cite{Watkins:1989} and bandit problems ~\cite{Auer2002}, to trade-off
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

Let $l(\Delta^{'}_t)$ denote this time the number of jobs \textit{finished} during
the period $\Delta^{'}_t$. We define the cost of policy $P_i$ during period
$\Delta^{'}_t$ as:
%
\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta^{'}_t)}
  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta^{'}_t
  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
%
As one can note, the actual waiting time is used here, so that one dispenses
with the use of a simulator for efficiency considerations. This leads to a
faster and easier estimate of the cumulative waiting time, however only for the
policy that is used during $\Delta^{'}_t$. Note that there is as previously a bias
due to boundary effects in the measurement method, as we do include jobs that
had been started before $\Delta^{'}_t$, as well as disregard jobs that were
submitted during $\Delta^{'}_t$ but finish later.

%This also explains the difference
%between the definitions of $l(\Delta^{'}_t)$: in the previous selection strategies,
%one relies on a simulated estimate so that jobs need not be finished during
%$\Delta^{'}_t$, but only submitted; in contrast, to know their actual waiting time,
%jobs need be finished during $\Delta^{'}_t$.

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w_{\mbox{epsilon}}(\rightarrow T;P_i) =
  \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta^{'}_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta^{'}_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta^{'}_t)$ is necessary to ensure that policies
remain comparable over time. Not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small either to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

The selection of the policy to be used during the current time period
$\Delta_T$ is then based on the standard $\epsilon$-greedy exploration
strategy:

%
\begin{itemize}
    \label{alg:epsilon}
\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
%
\begin{equation}
\label{eq:select-greedy}
P = \mbox{argmin}_{P_i, 1 \le i \le 12} w_{\mbox{epsilon}}(\rightarrow T;P_i)
\end{equation}
%
\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 12$, at random (exploration mode).
\end{itemize}

Figure~\ref{fig:bandit} illustrates this process. Several studies study the
decay of $\epsilon$ over time, the exploration being less important once the
estimates for the different policies are reliable ~\cite{Tokic:2010}. This
strategy has however not been beneficial in our case. We believe this is due to
the properties of the traces we use, as explained in
Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
outperformed in terms of regret by other classical algorithms such as the UCB
or EXP family of algorithms (see ~\cite{bubnow}). See ~\cite{bubnow} for a precise definition and
thorough treatment of this subject. These algorithms however all rely on
pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
in advance. While the usual doubling tricks can be devised to go around
sequential issues of scale with good asymptotical performance, they ruin
performance in practice. Accordingly, we only perform experiments using the
simple $\epsilon$-greedy method.

\begin{figure*}[]
  \centering
  \input{dia/_build/bandit.tex}

  \caption{Illustration of the bandit-based strategy. Every week,
  the wokload from the past week is used to run simulations using every
  reordering policy. The "model" $w$ is updated according
  to~\eqref{eq:tot-cost-greedy} and the policy for the next week is chosen
  according to~\eqref{alg:epsilon}. We denote by $P_{\mbox{bandit},t}$ the 
  policy chosen by the bandit at week $t$.
  }.

  \label{fig:bandit}
\end{figure*}

\section{Experiments}
\label{sec:experiments}

This section compares the different approaches to policy selection via a
comprehensive experimental validation.  Subsection~\ref{sub:protocol} outlines
our experimental protocol and Subsection~\ref{sub:results} contains the
experimental results and their discussion.

\subsection{Experimental Protocol}
\label{sub:protocol}

The evaluation of computer systems performance is a complex
task. Here, the two main difficulties are choosing a simulation
model and taking into account the variability in the average waiting times.
This section presents our approach, which is based on lightweight simulation
and trace resampling.

\subsubsection{Traces}
\label{sub:traces}

\begin{table}[]
  \centering
  \ra{1.3}
  \caption{Workload logs used in the simulations.}
  \label{tab:logs}
  \begin{tabular}{@{}lrrrr@{}}
    \hline
    Name          & Year & \# MaxProcs & \# Jobs & Duration\\
    \hline
    KTH-SP2       & 1996 & 100         & 28k     & 11 Months\\
    CTC-SP2       & 1996 & 338         & 77k     & 11 Months\\
    SDSC-SP2      & 2000 & 128         & 59k     & 24 Months\\
    SDSC-BLUE     & 2003 & 1,152       & 243k    & 32 Months\\
    ANL-Intrepid  & 2009 & 163,840     & 68k     & 9  Months\\
    CEA-Curie     & 2012 & 80,640      & 312k    & 3  Months\\
    Unilu-Gaia    & 2014 & 2,004       & 51k     & 4  Months\\
    \hline
  \end{tabular}
\end{table}

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. These platforms are all targeted by this
work, since they all are homogenous systems that run a batch scheduler.
Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
Parallel Workload Archive ~\cite{Feitelson20142967}. We use the
``cleaned'' ~\cite{pwa-clean} version of the logs as per the archive. We impose
an additional filtering step to the workloads\footnote{This filtering step is
available in the reproducible workflow ~\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

\begin{enumerate}

  \item If the number of allocated or requested cores of a job  exceeds the
    size of the machine, we remove the job;

  \item If the number of allocated or requested cores is negative, we use the
    available positive value as the request. If both are negative (no data is
    available), we remove the job;

  \item If the runtime or submission time is negative (no data is available),
    we remove the job.

\end{enumerate}

\subsubsection{Simulator}

The choice of a simulator is a critical part of experimental validation that
raises a trade-off between precision and runtime. High precision can be
obtained by carefully modeling the platform and its network topology, or
extracting information from the jobs from their sources or post-mortem logs.
This is the approach used by high-fidelity batch scheduling simulators such as
BatSim ~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without addressing the allocation
problem. Moreover, the experimental protocol used here requires many
simulation runs. Therefore, we set the precision/runtime trade-off at the
point which minimizes simulation runtimes. We discard all topological
information relative to the platform and use the job processing times of the
original workloads. In this setup, the processors are considered to be
undistinguishable from each other, and jobs can be discontinuously mapped to
any available processor on the system. During this work, we developed a
lightweight simulator ~\cite{ocst} and now release it to the community along with
this article. See the reproducibility paragraph below for more information.
With this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in
33 seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a comparison, the Batsim simulator with network communication
modeling disabled and no resource contiguity takes more than 7 hours to replay
the same trace on the same hardware. The complete set of simulations presented
below is executed on a single Dell PowerEdge R730 in 22 hours, including input
preparation and statistical analysis code.

\subsubsection{Resampling}
\label{ssub:resampling}

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics ~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem ~\cite{flurries} that motivated decades of research in workload
modeling ~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in ~\cite{feitresampling}, an approach that attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, as illustrated in~\cite{jsspp17},
it is reasonable
to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. In our experiments, we use stationary resampled data, 
and therefore no decay is necessary: $\lambda$ here is set to $1$.

%Thus,
%computing an average over the sampled traces at a given time can be related to
%a time average of an underlying Markov process, which converges to the true
%quantity thanks to the ergodic theorem. Moreover, the job submission process
%has no long range correlation making it look like an independent and
%identically distributed process when considering sufficiently time spaced
%weeks.

\subsubsection{Experimental Validation}

For each of the 7 workloads
considered, we use the original data to resample 60 traces, each with a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours. The choice of this default
stems from the analysis developed in ~\cite{jsspp17}. Essentially, this value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

\begin{itemize}

  \item All fixed $P_i$ for $1 \leq i \leq 12$ from the search set defined in
    Subsection ~\ref{subsec:policies}. This is done for two reasons. First, it allows to
    see which policy works well for which platform. Second, it allows to
    compare how well strategies used in this work with the best
    \textit{a-posteriori} known policy;

  \item For all the above strategies based on exact simulation (referred to as
    ``Simulated feedback'' in the remainder), on noisy simulation (referred to
    as ``Noisy Simulated Feedback'') and on $\epsilon$-greedy exploration
    (referred to as ``Bandit Feedback''), we experiment with two policy
    periods: $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This
    choice is a consequence of the natural rhythm of human activity: Daily or
    weekly simulation periods seem to ease comparing of performance between two
    periods.  Indeed, arbitrary period values would needlessly increase the
    variability of the cost metric. For the bandit feedback strategy, $\lambda$
    is set to $1$. This means that we do not use any decay in the estimation of
    the overall cost. Choosing to eliminate decay in this experimental
    validation is natural since our resampled data is in a stationary regime
    \textit{by construction}, as explained in~\ref{ssub:resampling}. The
    $\epsilon$ hyperparameter will be set at a value of $0.1$ for all traces, which is the
    result of a leave-one out majority vote selection process on
    $[0.1,0.3,0.5,0.7,0.9]$. In this process, one trace among the 7
    from Table~\ref{tab:logs} is taken out, and the most frequent best
    hyperparameter value is used. Table~\ref{tab:hyperparam} outlines the
    result of this cross-validation. Assuming workload traces are i.i.d, this
    ensures that the reported results are fair with respect to the situation
    where the reader of the study would implement the strategy directly.

  \item Lastly, an additional ``Random'' strategy is used as a reference point,
    with, as before, $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$.
    This strategy corresponds to choosing $P_i$ uniformly at random in
    $\left\{P_1 , \cdots, P_{12} \right\}$ during every period $\Delta_t$.

\end{itemize}

\begin{table}[]
  \centering
  \ra{1.3}
  \caption{Hyper-parameter leave-one-out selection for $\epsilon$.}
  \label{tab:hyperparam}
  \begin{tabular}{@{}lr@{}}
    \hline
    Name          & best $\epsilon$\\
    \hline
    KTH-SP2       & 0.1 \\
    CTC-SP2       & 0.1 \\
    SDSC-SP2      & 0.1 \\
    SDSC-BLUE     & 0.1 \\
    ANL-Intrepid  & 0.3 \\
    CEA-Curie     & 0.3 \\
    Unilu-Gaia    & 0.1 \\
    \hline
  \end{tabular}
\end{table}

The experimental validation uses of 7 traces $\times$ 60 resamples $\times$
(12+2+2+2) algorithms, which correspond to 12600 scheduling simulations spanning
two years. This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

\subsection{Replicability}

We publish a declarative workflow ~\cite{repro} based on our
simulator ~\cite{ocst} that runs experiments and generates all the figures from
this paper.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach ~\cite{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

All the dependencies of our
experiments (including our own code, dependencies for data processing and
visualization and workflow engine) are automatically managed by Nix up to the
actual execution of the code.
The Nix package manager is available at \url{https://nixos.org/nix}.

The archive containing the experimental workflow can be obtained at:
\begin{lstlisting}
https://github.com/freuk/obps
\end{lstlisting}

All the code associated with this work is released under the ISC ~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

\begin{lstlisting}
nix-build -A banditSelection
\end{lstlisting}

The experiments in this paper can run in less than 24 hours on a
moderately parallel host. More precisely, we had access to a Dell PowerEdge
R730 equipped with a total of 56 threads @2.4GHz each and 757 Gigabites of RAM. For
this reason, the experimental workflow was designed to be executed on
a single host.

All data from this paper were obtained from the Swf Parallel Workloads
Archive ~\cite{Feitelson20142967}.

 All experiments are tied together using zymake ~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments\footnote{The zymake system
 is also packaged by our nix expressions.}. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workflow is composed of the
 following steps:

 \begin{itemize}
   \item Data extraction from archives, data filtering, and data resampling.
     This is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{ssub:resampling}.

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. This simulator is made
     available under the ISC ~\cite{isc} license as a persistent zenodo
     archive at ~\cite{ocst}.

   \item Analysis of the results. This step runs R code that generates the
     figures presented in this paper.
 \end{itemize}

All resulting figures will be situated in the simlinked folder named
\lstinline[basicstyle=\ttfamily\color{blue}]|result| positionned at the root of
the archive.

\subsection{Results}
\label{sub:results}

\subsubsection{Analysis of Basic Policies}

\begin{table*}[]
  \centering
  \ra{1.3}

  \caption{Average total waiting time diminution of fixed policies with respect
  to EASY(FCFS). The precise definition of the value reported is provided in
Eq~\eqref{eq:cum}}

  %DELTA WEEK

  \label{tab:full}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & LCFS  & LPF            & SPF   & LQF            & SQF   & SEXP  & LEXP           & LRF            & SRF   & LAF            & SAF   \\
    \hline
    KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%          & +15\% \\
    CTC-SP2      & -40\% & -19\%          & -14\% & \textbf{-47\%} & +15\% & -36\% & -19\%          & -38\%          & -12\% & -44\%          & +22\% \\
    SDSC-SP2     & -8\%  & \textbf{-15\%} & +4\%  & -8\%           & +4\%  & -3\%  & \textbf{-15\%} & -3\%           & -10\% & -11\%          & +18\% \\
    SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%          & +5\%  \\
    CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%          & -17\% \\
    Unilu-Gaia   & -58\% & -33\%          & -10\% & \textbf{-62\%} & +16\% & -49\% & -27\%          & -56\%          & -16\% & -62\%          & +20\% \\
    \hline
  \end{tabular}
\end{table*}

We begin by drawing the attention of the reader to Table~\ref{tab:full} that
reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. Let
us denote the resampled traces of a given workload by $T_k$, where k is the
index of the resampled trace ($1 \le k \le 60$). Let the number of jobs in the
resampled trace $T_k$ be $l(T_k)$. Let $\bar{w}(T_k,P)$ be the cumulative
waiting time obtained by a simulation run for a strategy $P$:

\begin{equation}
  \bar{w}(T_k,P) = \sum_{j=1}^{l(T_k)} \text{Wait}_j^{P}
\end{equation}

The values appearing in Table~\ref{tab:full} are exactly:

\begin{equation} \label{eq:cum}
  \frac{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(P_i,40h))-
  \bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}{\sum_{k=1}^{k=60}\bar{w}(T_k,\text{EASY}(\text{FCFS},40h))}
\end{equation}

The best values are outlined in bold in the table. These values give the best
attainable performance in the search set defined in~\ref{subsec:policies}.
Note that there seems to be some regularity; for instance the SAF policy is remarkably
inefficient and the LQF policy works very well. However there are traces for which LQF is not optimal. Moreover, these
are~\textit{a-posteriori} results, which means that we only know
\textit{post-factum} that this policy would have been better.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/full-UniLu-Ga.pdf}
  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 12$ on the
    UniLu-Gaia trace. The average is obtained by resampling the original trace
    60 times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling.}
  \label{fig:all}
\end{figure}

In the present work, we are studying adaptive policies that obtain a good
performance on any given system by selecting one of these policies.
Figure~\ref{fig:all} reports the evolution of the average cumulative waiting
time for the UniLu-Gaia trace. This figure shows how the average improvement
in cumulative waiting time evolves as a function of time. In other words, this
is the average curve obtained if one were to plot the cumulative sum of job
waiting times for every resampled trace, incrementing the sum when a job
finishes. Accordingly, the final values attained by every curve correspond
exactly to the values reported in Table~\ref{tab:full}. The dashed values
represent the 10th and 90th percentile of the values.

Observe that while the 10th and 90th percentiles of the improvement with
respect to EASY-FCFS are large, the values are still contained enough to make
definite statements about how these policies compare to the FCFS baseline.
This figure does not report the variability of the policies against one another,
and does not allow to compare them beyond their average performance.
The fact that the size of the percentile band is roughly half
the best attainable improvement was our original motivation for this work, as
one can expect to be able to distinguish between these policies in an
online manner.

\subsubsection{Analysis of Policy Selection Strategies}

\begin{figure*}[]
  \centering
  \includegraphics[scale=0.6]{figures/CTC-SP2.pdf}
  \includegraphics[scale=0.6]{figures/KTH-SP2.pdf}\\
  \includegraphics[scale=0.6]{figures/CEA-Curi.pdf}
  \includegraphics[scale=0.6]{figures/SDSC-BLU.pdf}\\
  \includegraphics[scale=0.6]{figures/SDSC-SP2.pdf}
  \includegraphics[scale=0.6]{figures/ANL-Intr.pdf}

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the Simulated Feedback, Noisy Simulated Feedback and Bandit Feedback
    policies. The average is obtained by resampling the original trace 60
    times. The dashed lines represent the 10th and 90th percentiles of the
  values across this resampling. Each figure is a different trace. This figure
is followed-up in figure~\ref{fig:follow} for the remaining UniLu-Gaia trace.}

  \label{fig:small}
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/UniLu-Ga.pdf}

  \caption{Figure~\ref{fig:small} continued, plot for the UniLu-Gaia log.}

  \label{fig:follow}
\end{figure}

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
results concerning the strategies used in this paper. These figures show the
average cumulative improvement over EASY(FCFS,40h) of the three strategies
we use in the same format as Figure~\ref{fig:all}. There are five
strategies displayed on the graph:

\begin{itemize}
  \item Random $P \in P_i, 1 \le k \le 60$. ($\Delta=$1 week)
  \item Bandit Feedback ($\Delta=$1 week)
  \item Bandit Feedback ($\Delta=$1 day)
  \item Simulated Feedback ($\Delta=$1 week)
  \item Noisy Simulated Feedback ($\Delta=$1 week)
\end{itemize}

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
    \hline
    Trace        & Best(hindsight) & Random &       & Full           &                & Noisy &       & Bandit & \\
    \hline                
    $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
    CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
    SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

Table~\ref{tab:strat} reports the average total waiting time performance
improvement over EASY(FCFS,40h) of all the strategies used along with the
random variant. We make the following observations.

\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
~\ref{subsec:policies} is better than using the FCFS order. This is due to the
known fact that the FCFS order is not optimal for optimizing the average
waiting time objective, and confirms that the search set is well chosen. The
value of the period length $\Delta$ has an impact on the performance of the
random policy. Indeed, shorter periods lead to better performance.

\textbf{The Simulated Feedback strategy consistently performs almost as well as the
best fixed policy.} The values attained by the exact simulation based strategy almost match
the highlighted values from Table~\ref{tab:full}. This corresponds to the curve
labeled ``Full'' in Figures~\ref{fig:small} and~\ref{fig:follow}. Moreover,
these average cumulative waiting time curves exhibit linearity (as opposed to
being concave), which further shows that the choice of the best policy can be
made rapidly. Note that the value of $\Delta$ has no impact over the
performance of this strategy.

\textbf{Noisy simulations can be used.} The performance attained by the noisy
variant almost matches that of the exact-simulation based strategy. This indicates
that it is not necessary to have access to a precise simulator. Rather, we may
hope that an imprecise but unbiased reporting of the performance is enough to
rapidly select a reordering policy.

\textbf{Full vs Bandit feedback.} As expected, the bandit strategy evolves
between the random strategy and the full feedback strategy. The bandit-based
strategy always fares better with a shorter period $\Delta$, recovering between
a quarter to the whole gap between the Random and Full Feedback strategies.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by the Noisy policy ($\Delta=$1 day,
  $\lambda = 1 $) as a function of time
    for the UniLu-Gaia trace. The average choice
  is obtained by resampling the original trace 60 times and aggregating by
  date. The legend excludes policies that are never used by
  the strategy.}

  \label{fig:mosn}
\end{figure}

\textbf{Stability of the policy selection.}
Figure~\ref{fig:mosn} gives insight regarding the behavior of the noisy
feedback policy with $\Delta=$ 1 week. This figure reports the
share of the policies chosen by the strategy on average as a function of time.
This is an area plot: The height of the colored region
represents the value. Moreover, the proportion of times each policy is
chosen by the strategy is aggregated across resampled traces.

While this policy obtains an excellent performance, the convergence to a fixed
policy is slow. This is not harmful \textit{per se} and just means that it is
difficult to distinguish between the best policies. Additionally, we observe that
the two increasingly dominant policies, LAF and LQF, are the best overall policies.
This means that the estimation converges to the correct point.

\begin{figure}[]
  \centering
  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by the Bandit Feedback($\Delta=$1 day,
  $\epsilon = 0.5, \lambda = 1 $) policy as a function of time for the UniLu-Gaia
trace. The average choice is obtained by resampling the original trace 60 times
and aggregating by date. The legend excludes policies that are never used by
  the strategy.}

  \label{fig:mosb}
\end{figure}

Figure~\ref{fig:mosb} is the analogue of Figure~\ref{fig:mosn}, this time for
the Bandit Feedback policy with $\Delta$ = 1 week. Observe that here the choice
of policy is much less stable. The convergence is extremely slow, which is
expected given the variability of the reward we give to the bandit algorithm.

\section{Conclusion}
\label{sec:ccl}

The scheduling of parallel jobs on a given HPC platform is a very hard
optimization problem with many uncertain parameters. Even though these
uncertainties could be reduced, determining efficient strategies remains
difficult.

In this work, we presented a new way of addressing this problem. The idea was
to look directly at the scheduling process instead of trying to change its
parameters.

More precisely, we showed that it is worth learning on the scheduling process
itself. Indeed, reordering the submission queues under EASY-Backfilling leads
to considerable gains in performance.

This approach was used in two methods: a method based on a simulator and a
method based on a multi-armed bandit algorithm. The first one provides very
good results, reducing the average waiting times of the baseline FCFS ordering
policy by 11\% to 60\%. The second one is slightly less efficient with an
improvement factor of 8\% to 46\%. However, the bandit-based method is easier
to use and cheaper to run.

There are two directions to extend this work. First, the question of
the existence of a more efficient selection strategy for bandit feedback for
these systems is not addressed by this work. Is a simulator really necessary?
%@Denis: a essayer.. 1 mois de travail pour tester?
Second, one can't help but wonder what performance could be achieved by extending
the search space to a wider class of policies. This means studying learning performance using some
function approximator.
%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
As both these directions are data-bound, they will require careful application of
trace resampling techniques to be successful.

\section*{Acknowledgements}

Authors are sorted in alphabetical order. The authors would like to warmly
thank Pierre Neyron for help with computational experiments, as well as Michael
Mercier, Salah Zrigui, Bruno Raffin and Theophile Terraz for their comments and
Arnaud Legrand for his work on reproducibility. We gracefully thank the
contributors of the Parallel Workloads Archive, Victor Hazlewood (SDSC SP2),
Travis Earheart and Nancy Wilkins-Diehr (SDSC Blue), Lars Malinowsky (KTH SP2),
Dan Dwyer and Steve Hotovy (CTC SP2), Joseph Emeras (CEA Curie and UniLu Gaia),
Susan Coghlan, Narayan Desay, Wei Tang (ANL Intrepid), and of course Dror
Feitelson. The Metacentrum workload log was graciously provided by the Czech
National Grid Infrastructure MetaCentrum. This work has been partially
supported by the LabEx
PERSYVAL-Lab(ANR-11-LABX-0025-01) funded by
the French program Investissement d'avenir.
 Experiments presented in this paper were carried out using the Grid'5000
 testbed. Grid'5000 is supported by a scientific interest group hosted by Inria
 and including CNRS, RENATER and several Universities as well as other
 organizations\footnote{https://www.grid5000.fr}. Access to the experimental
 machine(s) used in this paper was gracefully granted by research teams from
 LIG\footnote{http://www.liglab.fr} and Inria\footnote{http://www.inria.fr}.

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}