### 2018-01-11 paper/revision.tex
revision letter

\documentclass[]{article}

\usepackage{listings}
\usepackage{fullpage}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

\usepackage{amssymb}
\usepackage{tabulary}

\usepackage{etoolbox}

\usepackage{algorithm}
\usepackage{algorithmic}

\title{Major revision comments.}

\begin{document}

\maketitle

\section{Reviewer 1}

\textbf{The manuscript describes an experimental study to modify a scheduling
approach common in many large parallel computer systems. The authors propose to
replace the well-established First-Come-First-Serve approach combined  with
EASY-backfilling by a strategy using reordering of the queue. To prevent job
starvation, the authors resort to FCFS for a job as soon as the waiting time of
this job exceeds a threshold. Although this threshold approach is well known,
the authors do not provide any references of previous publications using this
approach. }

\textbf{The reviewer sees significant shortcomings in the experimental design
of the manuscript. In general, computation experiments should follow the
guidelines of experiments in natural sciences, that is, an experiment has the
goal to prove or disprove a given hypothesis. The authors do not describe such
hypothesis nor can the reviewer extract it from the manuscript (except for a
trivial hypothesis: we can beat EASY-Backfilling for some traces). }

\textbf{Instead, the authors present a simple trial-and error study that
exploits some parts of the problem space without justifying the restriction of
the problem space. For instance, the authors suggest 12 reordering policies
(Section 4.1) with the justification of the selecting a search space "as
semantically diverse as possible" instead of carefully arguing which reordering
policy has the potential to improve performance. For the scheduling objective,
the authors use "the waiting times of jobs, which is one of the more commonly
used objectives" although they state before that a "system administrator may
use one or multiple cost metric(s)". Although this approach is the opposite of
the approach used for the reordering policies, the authors do not justify the
design decisions of their study. }

\textbf{Similarly, the authors simply use 40h as threshold and simply state
that the "choice of this default stems from the analysis developed in" a former
manuscript of the authors that is so far unpublished. Such approach is simply
unacceptable in view of the reviewer.}

This is indeed a mistake on our part which made the way to the final version,
this is now corrected. The article is "Tuning Backfilling Queues" from JSSPP17.\\

\textbf{The authors use 7 traces recorded over a time span of 18 years. They do
not discuss whether these traces a comparable with respect to their study
despite large differences in the number of processors, jobs and trace lengths.}

\textbf{In their algorithmic description, the authors include a decaying factor
$\lambda$ although they later simply state "this will not be studied in the
experiments where $\lambda=1$ and is left for future research". Therefore, the
authors indicate that even in their own opinion, research regarding this
algorithm is not complete but the presented findings are only a first step. The
reviewer feels that such first step may be helpful in a workshop to discuss
directions to complete such study but not for reporting the results of a study
in a journal. }

\textbf{Altogether, the reviewer believes that an experimental study with so
many design problems is not relevant to the readers of TDPS. }

\textbf{The reviewer does not see how the authors can overcome these
shortcomings even in a major revision and therefore suggest rejection of the
manuscript.}

\textbf{In addition, the reviewer sees several minor problems in the manuscript
that the authors can correct in a major revision. The reviewer only briefly
describes such problems in the following:}

\textbf{Abstract: Any reader of TPDS can expect to understand the abstract
without having to use additional literature. Therefore, the abstract should
only include expressions that are common knowledge: it is at least doubtful
that every reader understands the used abbreviations and expressions like
"Epsilon-greedy multi armed bandit algorithm". }

\textbf{Previous publications explain that studies using traces suffer from
fill-ins in the starting period and from draining in the final period. The
authors mention this problem but do not discuss their impact and strategies to
avoid these problems although other publications present such strategies.
Previous studies have described that the KTH trace is hardly comparable with
other traces due to a specific acceptance policy. The authors do not discuss
this problem. The authors use either 1 week or 1 day feedback justifying their
choices with the periodicity of the traces. Although such approach is
reasonable, it is a hypothesis and requires validation by using other feedback
periods as well and comparing the results. }

\textbf{There are several publications discussing the disadvantages in using
traces and suggesting workload models instead. The authors do not consider
these studies in their manuscript without explaining why.}

We leave this discussion to other works, in particular, we suggest the EuroPar
article 'Workload Resampling for Performance Evaluation of Parallel Job
Schedulers'.\\

\textbf{Problems of the introduction explaining the goal of the study are
likely related to the already mentioned design problems of the study. The same
holds for the organization of the manuscript.}

\textbf{The authors can improve readability of the manuscript by explaining the
underlying algorithmic approach in some more detail.}

We added two figures to illustrate the algorithms.\\

\textbf{Since the journal provides a platform for supplemental material, the
reviewer does not understand why readers must contact the authors by email to
obtain data to reproduce the study.  }

As previously concerning the missing reference, this was inadvertent on our
part! This is fixed in the major revision. Apologies for this error.

\section{Reviewer 2}

\textbf{I basically liked this paper and support its publication.  The
following comments make some suggestions for possible improvements.}

\textbf{Title: I would suggest to remove the "(bandit)" from the title, and
instead use something like "Online Queue-Ordering Policy Selection for
EASY-Backfilling" -- emphasize what you are doing, not how.}

The need for more careful framing of the goals of the study was stressed by all
reviewers. Accordingly, we made the scope of the study more clear and changed
the title, to "Online Tuning of EASY-Backfilling using Queue Reordering
Policies."\\

\textbf{Related work: additional papers you may want to look at are}
\begin{enumerate}

  \item \textbf{"Exploring portfolio scheduling for long-term execution of
    scientific workloads in IaaS clouds", from SC'13, suggests how to select
    from a large portfolio of possible schedulers;}

  \item \textbf{"Improving and stabilizing parallel computer performance using
    adaptive backfilling", from IPDPS 2005, which also suggests online
    selection from a set of 4 alternative schedulers;}

  \item \textbf{"Selective reservation Strategies for Backfill Job Scheduling",
    from JSSPP 2002, which suggests online decisions of whether to make
    reservations for skipped jobs - similar effect to your thresholding
    scheme;}

  \item \textbf{"Self-tuning systems", from IEEE Software 1999, which is about
    simulating performance with local workloads to select systems parameter
    values.}

\end{enumerate}
Thank you for your suggestions, we had not come across these papers! There are indeed very related to our effort.

\textbf{Section 2.2 - could be good to give a short description of how a basic
    version of such an algorithm works, instead of just noting a list of
    versions.}

We did add a basic description of the setting in the description.\\

\textbf{Section 3.1 notation - intuitively p could represent processors just as
well as processing time, so I would suggest $t_i$ for time and $p_i$ for
processors.  Similarly, in Section 4.1 I for one would find the names more
intuitive with SJF/LJF instead of SPF/LPF (shortest/longes job first) SPF/LPF
instead of SQF/LQF (smallest/largest processors first) expansion factor is
often called slowdown area is often called total resources}

We agree that $p_i$ is not very intuitive, but for lack of a standard we chose
to use notation from scheduling theory.  In face of the many possible acronyms,
we chose to rather be consistent with the $p_i$ and $q_i$ notation - SPF ist
smallest $p_i$ first and SQF is smallest $q_i$ first..  We added the remark
concerning the total resources and the expansion factor. This notation comes
from the "Characterization of backfilling strategies for parallel job
scheduling" article, where the authors call Expansion Factor the value of the
slowdown where the upper bound on the running-time is used.\\

\textbf{Section 5.1 exact simulation: I understand why you divide the time into
    periods as part of online tuning. But I disagree with doing so also in the
    simulations - as you yourselves say, it just causes problems with boundary
    effects.  And you can discount more distant jobs without using periods.}

We agree that this design choice can be discussed. Our decision to divide
makes the study more realistic with respect to the cases where simulations are
expensive(which is most large systems).  Indeed, the more realistic simulations
are, the more time they take. For example, precise simulation frameworks such
as SimGrid/Batsim have high runtimes, and simulating the history of the whole
system every period is prohibitive if the network topology and/or multiple
resource constraints are used. As for discounting - it can indeed be used
in both cases.\\

\textbf{Section 5.2 - adding noise to the output rather than to the input
    workload doesn't make sense in my opinion. And the result that it gave the
    same results is totally expected - you add random noise with mean 0 and
    then fined that the average didn't change.}

We do agree that this does not change the estimate - the point of adding 
this noise was to show that the sample size is not prohibitive in practice.
Indeed, we learn the policy almost as fast with this noise, which
motivates the use of the technique in real-world cases. The point of this noise
is really to compare the difference in the shapes of the two green curves
in Figure 3.\\

\textbf{Section 6.1.1 traces - I think negative values are just -1 meaning no data is available.}

You are right - and we made this point clear in the text.\\

\textbf{Section 6.3.1 - I think you can say more about the post-factum results.  Even
if it is not your focus, it is still interesting that for example ordering
by size is more beneficial than by time, and that it should be
largest-first (and the SJF-like scheme is not good).}

Doing this would drown the point in the author's opinion. We agree it is an
interesting point - we now refer to our previous publication 'Tuning
Backfilling Queues'(JSSPP17), which was inadvertently ommited from the related
works of the first submission.\\

  \textbf{Section 6.3.2 FCFS vs. Random: you justify the result be a conjecture that jobs do not pile up in the queue.  you should verify this by counting the jobs int he queue in the simulation.}

We aggree that this conjecture is not warranted - we removed it in the revision.
We do have plans to explore this aspect for a future publication where the
discussion will be more appropriate.\\

\textbf{Table 3: add columns with best static result for comparison.}

Done, thanks you for the suggestion.\\

\textbf{Description of Figure 6 results: this is not slow convergence, it is no
convergence.  and using epsilon=0.5 is much too high: it means that half of the
time you are choosing randomly, and that's why all of the possible approaches
are used approximately the same fraction of the time. What you want is to
initially use them all, but then reduce epsilon with time till you reach some
reasonable value like random exploration in 1 of 10 or 1 of 20 periods.}

We changed the experimental protocol to use leave-one-out cross validation,
which did improve the bandit results. The selected exploration rate was 0.1 for
all traces, although it is not the best choice for all of them. (see the new
Table 2) The choice of a fixed exploration rate is one of simplicity - while
decaying exploration does provide better performance, it is also harder to
tune.\\

\textbf{Reference [4] is often referred to and gives important background, but
is not available. NOTE: this journal does not really do blind review. Your
names are given on the first page.}

This is corrected in the major revision - we inadvertently ommited this
when preparing the submission. The reference is 'Tuning Backfilling Queues'(JSSPP17).\\

\textbf{Typos and usage [...]}

We incorporated these changes and thank the reviewer for their detailed read.\\

\section{Reviewer 3}

\textbf{I think that the current experiments are not sufficient to show the advantage of the proposed strategy.
The authors compare the performance of the proposed strategy with the baseline performance derived by EASY Backfilling. The performance of EASY Backfilling is good metric as the baseline, but comparison with EASY Backfilling is not enough to discuss the advantage of the proposed strategy. A lot of work about tuned Backfilling has been presented as discussed in Sec. 2. The author should discuss the performance improvement compared with the existing tuned backfilling strategies.
The experimental results in Table 2 show that LAF and LQF show best performance among 12 reordering policies. This indicates that we can expect significant performance improvement by using static strategies, Backfilling with LAF or that with LQF, compared with EASY Backfilling. If the performance gap between the proposed strategy and the static strategies are small, the static strategies may be preferable because they are less complicated and run with smaller scheduling overhead. A comparison between the proposed strategy and the static strategies is needed to show the advantage of the proposed strategy.}

We do agree with the reviewer that in order to really motivate the use of
bandit/resimulation aproaches, one should compare the performance of static
policies, say using leave-one-trace-out cross validation, to the performance of
the adaptive variants, using leave-one-trace-out cross validation for their
hyperparameters.
However, this work does not aim to underline the necessity of the bandit
approach (the title has been changed and the introduction modified in the major
revision to reflect this better). Rather, we try to show its advantages,
feasibility and limitations. 
There are two main reasons for not trying to be too demonstrative of the superiority
of the adaptive approach using our hypotheses.

First, this study uses a limited amount of workload logs. In consequence, the
process of comparing the adaptive (hyperparameter-cross-validated) choice to
the static (cross-validated in the set of fixed policies) choice would not be
really conclusive.  Indeed, the more workload logs are presented, the more
diversity is present and the more adaptation is necessary.
Second, we use a simplistic simulation model. In this model, we do make
machines *more* similar to each other by neglecting topological differences. It
can be reasonably expected that differences between systems should *increase*
rather than decrease, when moving away from a simplistic model to the real
systems.
We modified the present experiments to use leave-one-out cross
validation in the hyperparameter selection, and made the reasons for this
choice of experimental protocol clearer in the text.
We are preparing a subsequent study using an approach closer to what you
describe, in which we greatly increase the number of workload logs used.\\

\textbf{The authors assume +/-20% noise in the simulation running in the
proposed strategy. However, the reason why this assumption is suitable is not
presented. The authors should discuss the noise of 20% is suitable to represent
real world jobs.}

A justification was indeed missing, this value is taken as to be sufficiently
larger than values previously reported, see e.g. Batsim: a Realistic
Language-Independent Resources and Jobs Management Systems Simulator which
reports 2 percent variations of the simulated metric compared to the metric on
the real systems. We added this remark to the text.\\

\textbf{Minor comments for improving presentation}

We incorporated these changes and thank the reviewer for their detailed read.\\

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}